Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-02-19 Multiresolution_framework/ltx/manuscript.tex
minor edit to result section
minor edits to result section

The EM algorithm was allowed to run until a maximum number of 20 iterations was reached, though typically the change in the transition matrix Frobenius norm dropped below $10^{-6}$ after less than 10 iterations. The rate of convergence of the EM based algorithm is depicted in Fig.~\ref{fig:MRA-Convergence}. 
\begin{figure}[t]
 	\centering
 		\includegraphics[scale=1]{./Graph/fig8.pdf}
 		\caption{{\bf Convergence of the EM based algorithm}. Representative plot of
the $\mathbf A(\boldsymbol\theta)$ Frobenius norm vs iterations of EM based algorithm. The change in the stopping criterion falls below $10^{-6}$ after less than 10 iterations.} 
\label{fig:MRA-Convergence}  
 \end{figure}     
% The spatial cutoff frequency of the observed field was used to specify the level of decomposition required to represent the field using the wavelets. Using an oversampling parameter of 5, the cutoff spatial frequency was $\nu_{cy} \approx 6.9 $ cycles/mm, obtained by averaging the spectral power of the spatial frequency (over time) from the observations \cite{Scerri2009}. Therefore, wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field, yielding $n_x = 131$ states. 
The actual connectivity kernel and the decomposition is plotted in Fig.~\ref{fig:KernelEstimate}(A). No assumptions where made about the shape of the kernel. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The large standard deviation is due to the high number of parameters need to be estimated. The small error in the estimate is attributed to the MRA of the system used to form the estimator. Fig.~\ref{fig:KernelEstimate}(B) and (C) illustrate the kernel scaling and wavelet functions, respectively. A total number of 25 basis functions were used to reconstruct the kernel, comprising 13 and 12 scaling and wavelet functions at $j=1$, respectively. 
\begin{figure}[!t]  
 \centering
\includegraphics{./Graph/fig9.pdf}
\caption{ {\bf Connectivity kernel estimate}. (\textbf{A}) The actual kernel and its components are shown by solid and dashed black lines, respectively. The mean kernel estimates over 200 realizations and confidence interval are shown by the red line and red shaded region ($\pm2$ std.). (\textbf{B},\textbf{C}) Kernel scaling and wavelets functions, respectively, at $j=1$ used in the estimation algorithm.}
\label{fig:KernelEstimate}
\end{figure}

@@

To study the performance of different models with various MRA capabilities, a single realization was obtained using Eqs.~\eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation} under the same circumstances of the previous experiment. The generated data set was then employed to estimate the underlying field and the connectivity kernel parameters using different state-space models, $j=0$ up to $j=4$. The RMSE of the field estimation for different models is shown in Fig.~\ref{fig:RMSE}. By increasing the value of $j$, error in the estimation decreases and converges to a steady value of 0.71, confirming that the model with $j=3$ is in fact adequate to capture the dynamics of the underlying field. The spatiotemporal characteristics of the reconstructed fields (A-E) along with the actual field (F)  are shown in Fig.~\ref{fig:FieldEstimation}. The experiment was performed several times and the results were consistent over each run.

>>>>

The EM algorithm was allowed to run until a maximum number of 20 iterations was reached, though typically the change in the transition matrix Frobenius norm dropped below $10^{-6}$ after less than 10 iterations. The rate of convergence of the EM based algorithm is depicted in Fig.~\ref{fig:MRA-Convergence}. 
\begin{figure}[t]
 	\centering
 		\includegraphics[scale=1]{./Graph/fig8.pdf}
 		\caption{{\bf Convergence of the EM based algorithm}. Representative plot of
the $\mathbf A(\boldsymbol\theta)$ Frobenius norm vs iterations of EM based algorithm. The change in the stopping criterion falls below $10^{-6}$ after less than 10 iterations.} 
\label{fig:MRA-Convergence}  
 \end{figure}     
% The spatial cutoff frequency of the observed field was used to specify the level of decomposition required to represent the field using the wavelets. Using an oversampling parameter of 5, the cutoff spatial frequency was $\nu_{cy} \approx 6.9 $ cycles/mm, obtained by averaging the spectral power of the spatial frequency (over time) from the observations \cite{Scerri2009}. Therefore, wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field, yielding $n_x = 131$ states. 
The actual connectivity kernel and the decomposition is plotted in Fig.~\ref{fig:KernelEstimate}(A). No assumptions where made about the shape of the kernel. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The large standard deviation is due to the high number of parameters need to be estimated. The small error in the estimate is attributed to the MRA of the system used to form the estimator. Figs.~\ref{fig:KernelEstimate}(B) and (C) illustrate the kernel scaling and wavelet functions, respectively. A total number of 25 basis functions were used to reconstruct the kernel, comprising 13 and 12 scaling and wavelet functions at $j=1$, respectively. 
\begin{figure}[!t]  
 \centering
\includegraphics{./Graph/fig9.pdf}
\caption{ {\bf Connectivity kernel estimate}. (\textbf{A}) The actual kernel and its components are shown by solid and dashed black lines, respectively. The mean kernel estimates over 200 realizations and confidence interval are shown by the red line and red shaded region ($\pm2$ std.). (\textbf{B},\textbf{C}) Kernel scaling and wavelets functions, respectively, at $j=1$ used in the estimation algorithm.}
\label{fig:KernelEstimate}
\end{figure}

@@

To study the performance of different models with various MRA capabilities, a single realization was obtained using Eqs.~\eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation} under the same circumstances of the previous experiment. The generated data set was then employed to estimate the underlying field and the connectivity kernel parameters using different state-space models, $j=0$ up to $j=4$. The RMSE of the field estimation for different models is shown in Fig.~\ref{fig:RMSE}. By increasing the value of $j$, error in the estimation decreases and converges to a steady value of 0.71, confirming that the model with $j=3$ (used in the first experiment) is in fact adequate to capture the dynamics of the underlying field. The spatiotemporal characteristics of the reconstructed fields (A-E) along with the actual field (F)  are shown in Fig.~\ref{fig:FieldEstimation}. The experiment was performed several times and the results were consistent over each run.
